Performance benefits of Intel Optane DC storage for the parallel processing of large neuroimaging data

### 2019-12-10 paper.tex
added updated scripts and text

\begin{abstract}
    Open-access neuroimaging datasets have reached up to $\SI{}{\peta\byte}$ of data, and
    continue to grow. The ability to leverage the entirety of these datasets are
    limited to a restricted number of labs with both the capacity and infrastructure
    to process the data. Whereas BigData engines have significantly reduced
    application performance penalties with respect to data movement, their applied
    strategies are not necessarily practical within neuroimaging workflows where
    intermediary results may need to be materialized to shared storage for post-processing
    analysis. In this paper we evaluate the performance advantage brought by Intel
    Optane DC persistent memory storage for the processing of large neuroimaging 
    datasets using the two available configurations modes: Memory Mode and App Direct Mode.
    We employ a synthetic algorithm on the $\SI{76}{\gibi\byte}$ \bigbrain
    and a real neuroimaging application on the Consortium for Reliability and Reproducibility (CoRR)
    dataset using 25 and 96 parallel
    processes in both cases. Our results show that performance of all available storage
    devices (e.g. tmpfs, local disk and Isilon) is somewhat deteriorated by memory mode
    when compared to App Direct Mode. In contrast, Optane in App Direct Mode was
    found to be slower than local disk, when the load on the bus was high due to sheer
    amount of concurrent processes and short task duration. It is believed that
    Optane would likely be useful in the accessing of typically larger-than-memory
    data in a limited-parallelism manner, such as with visualization applications.
    Further experiments with larger datasets remain to be performed.

\end{abstract}

@@

The server used for the experiments consisted of 12 \SI{64}{\gibi\byte} DRAM devices,
resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12 \SI{256}{\giga\byte} Intel Optane DC Memory
devices, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC Memory. The theoretical maximum bandwidth
of DRAM is estimated to be around \SI{275}{\gibi\byte}~/s whereas for Optane, it is
estimated to be around \SI{250}{\gibi\byte}~/s.
Other storage devices included a \SI{240}{\gibi\byte} Micron SATA SSD, of which
only \SI{149}{\gibi\byte} where available for the experiments, and a \SI{628}{\tebi\byte} shared Dell EMC Isilon cluster. The local disk
was the only storage device set up as a writeback device. Both Isilon and Optane in
App Direct mode were configured as write-through devices, meaning they did not leverage
DRAM as a cache for writes. This was achieved in Optane by configuring the xfs filesystem with
DAX. All storage devices were benchmarked using the following \href{https://github.com/ValHayot/paper-memory-storage/scripts/bench_disks.sh}{script}.
The result of the benchmarks can be seen in Table \ref{table:bandwidths}.

@@

@@

\subsection{40~$\mu$m \bigbrain incrementation}
\subsubsection{25 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_25cpu.pdf}
    \captionsetup{width=\columnwidth}
    \caption{Makespan of the \bigbrain Incrementation application using 25 
             processes on the 125 blocks on the 40~$\mu$m image given different
             storage configurations. Anticipated read and write duration was 
             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
             to Eq.~\ref{eq:makespan}. 3 
             repetitions were performed. *Local SSD did not complete the
    writing of the last few blocks due to storage limitations.}\label{fig:makespan-25cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/violin-read-20bb_25cpu.pdf}
    \captionsetup{width=\columnwidth}
    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 2. 3 repetitions were performed.}\label{fig:stacked-25cpus}
\end{figure}
\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567709282-tmpfsAD_1it_25cpus_120delay_withpc-2.pdf}
    \caption{tmpfs}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567705458-optaneAD_1it_25cpus_120delay_withpc-2.pdf}
    \caption{Optane}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567704324-localAD_1it_25cpus_120delay_withpc-2.pdf}
    \caption{local disk}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567707833-isilonAD_1it_25cpus_120delay_withpc-2.pdf}
    \caption{Isilon}\label{fig:gantt25isilon}
\end{subfigure}
\caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes and a sleep delay of 120s}\label{fig:gantt25}
\end{figure*}

@@

When observing the total task breakdown (Figure~\ref{fig:stacked-25cpus}), it is
apparent that the differences between the makespans arise from difference in incrementation
time between all devices except Isilon. In addition, while tmpfs and local disk have comparable write times, Optane
has a more significant write time difference. Whereas, Isilon has more significant read and
write time compared to all other devices, its incrementation time is comparable to that
of local disk. Further inspection of each storage device's Gantt chart (see Figure~\ref{fig:gantt25})
shows that increased duration of tasks stagger the execution of each individual tasks. This, in turn,
results in better performance towards the end of the execution, where tasks are significantly more staggered.
Only tmpfs appears to be able to execute each task efficiently with little staggering.
\subsubsection{96 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-96cpus-0delay-1it.pdf}
    \captionsetup{width=\columnwidth}
    \caption{Makespan of \bigbrain Incrementation using 96 processes on 125 blocks on the 40~$\mu$m image given different storage configurations. 5 repetitions were performed.}\label{fig:makespan-96cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-96cpus-0delay-1it.pdf}
    \captionsetup{width=\columnwidth}
    \caption{total read/increment/write breakdown of \bigbrain incrementation using 96 processes on 125 blocks on the 40~$\mu$m image given different storage configurations. 5 repetitions were performed.}\label{fig:stacked-96cpus}
\end{figure}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1568150082-local_1it_96cpus_0delay_npc-3.pdf}
    \caption{Memory Mode}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1567714252-localAD_1it_96cpus_0delay_npc-3.pdf}
    \caption{App Direct Mode}
\end{subfigure}
\caption{Gantt charts for local disk processing 125 blocks of the 40$\mu$m BigBrain using 96 processes with no delay}\label{fig:gantt96}
\end{figure*}

At 96 processes, it was expected that the application would not satisfy the inequality.
In Figure~\ref{fig:makespan-96cpus} we see that there is now a greater disparity between the
makespans of experiments executed in Memory and App Direct Mode, given the same storage. Local disk
in App Direct Mode is significantly faster than its Memory Mode counterpart. Local disk in App Direct
mode is even faster than Optane, by a factor of 1.7. Tmpfs was also found to perform better in App Direct Mode, although
the performance difference is not significant. Only Isilon's performance did not vary between the
two execution modes.

As with 25 processes, the main cause of performance variability appears to be the incrementation
time (Figure~\ref{fig:stacked-96cpus}). Local disk running in App Direct Mode has an average
incrementation time that is nearly 15~x faster than in Memory Mode. The incrementation time
for local disk is, on average, 6~x faster than on Optane.

Isilon in App Direct generally performed the same as in memory mode, however, the write
was almost 4000~seconds slower, on average, in App Direct Mode.

A further look at the Gantt charts between Memory Mode and App Direct Mode (Figure~\ref{fig:gantt96})
shows that the sole difference between the two modes is the duration of the incrementation step.
While both modes have an increased incrementation time, the incrementation duration in
Memory Mode is much more important than that of App Direct Mode.

@@

Although tmpfs offers the best performance in terms of makespan duration, Optane
is a close contender. While tmpfs is approximately 5x faster than Optane, Optane can be
as much as 12x faster than on Isilon (Figure~\ref{fig:20mksp25}). By using the results of 
our emulation, we can speculate that Optane will also be significantly faster than 
local disk in a real scenario. The real Optane execution is almost 3x faster than
the fastest local disk emulation configuration (i.e. Memory mode - Spark). As the
emulation results have been found to be faster than the real results, potentially as a side-effect
of increased accessed to cached data, it is expected
that the true makespan of the application running on local disk would be significantly
greater.

The read/increment/write breakdown (Figure~\ref{fig:20stackedp25}) shows that tmpfs
spends the least amount of time executing all the tasks, however, it spends more
time than Optane incrementing the data. Otherwise, tmpfs is nearly 470x faster with
reading and nearly 11x faster at writing than Optane. Despite not reaching memory
speeds, Optane is significantly faster than other persistent storage devices.
For instance, it can be almost up to 10x faster in reads than Isilon and almost 
9x faster in writes. It is important, however, to note that total read times in App
Direct mode for the Spark execution on Isilon are almost twice as fast as their 
equivalent Optane counterpart. On the other hand, this decrease read time is offset
by the write time which are 9x slower.

Emulated breakdown results also agree that the bulk of the differences in makespans
can be attributed to write time, with local disk writes being notably slower than
Optane. In fact, local disk write speeds nearly match that of Isilon. Isilon results
in a longer makespan due to increase read and incrementation time.
\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_25cpus.pdf}
    \caption{real}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-em-20bb_25cpus.pdf}
    \caption{emulated}
\end{subfigure}
\caption{Makespan of the incrementation algorithm processing the 20~$\mu$m BigBrain using
25 processes}\label{fig:20mksp25}
\end{figure*}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_25cpus.pdf}
    \caption{real}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-emulated-20bb_25cpus.pdf}
    \caption{emulated}
\end{subfigure}
\caption{Total read/increment/write breakdown of the  incrementation algorithm processing the 20~$\mu$m BigBrain using
25 processes}\label{fig:20stackedp25}
\end{figure*}

@@

With nearly a four-fold increase in parallelization, the makespan generally did
not improve other but a few 100s on Isilon. Overall, the general trend is similar
than with 25 cpus. As can be seen in Figure~\ref{fig:20mksp96}, Optane in Memory
mode is the fastest storage device. Furthermore, Optane in App Direct mode 
remains significantly faster than the other storage devices (e.g. Isilon).
However, emulated disk results appear to imply that the speedup difference
between Optane in App Direct mode and local disk will be smaller than at 25 cpus.

The vast majority of processing time appears to be spent on writing to the 
respective storage device with the exception of Optane in Memory Mode, which 
spends most of its time incrementing (Figure~\ref{fig:20stackedp96}). App Direct
mode executions appear to have reduced incrementation time coupled with increased
write times when compared to their memory mode equivalents. Furthermore, read
times appear to be increased for Isilon memory mode executions. These events are
not reflected in the emulated executions.

Emulated breakdowns demonstrate that local disk executions will remain slower 
than Optane in both App Direct and Memory modes. Results show that both 
incrementation and write times will be generally be longer on local disk than
on Optane.

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_96cpus.pdf}
    \caption{real}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-em-20bb_96cpus.pdf}
    \caption{emulated}
\end{subfigure}
\caption{Makespan of the incrementation algorithm processing the 20~$\mu$m BigBrain using
96 processes}\label{fig:20mksp96}
\end{figure*}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_96cpus.pdf}
    \caption{real}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-emulated-20bb_96cpus.pdf}
    \caption{emulated}
\end{subfigure}
\caption{Total read/increment/write breakdown of the  incrementation algorithm processing the 20~$\mu$m BigBrain using 96 processes}\label{fig:20stackedp96}
\end{figure*}
\subsection{BIDS App Example}
\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/25cores.pdf}
    \caption{25 processes}\label{fig:bm25}
\end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/96cores.pdf}
    \caption{96 processes}\label{fig:bm96}
\end{subfigure}
\caption{Makespan of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}
\end{figure*}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/25cores-sum.pdf}
    \caption{25 processes}\label{fig:bb25}
\end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/96cores-sum.pdf}
    \caption{96 processes}\label{fig:bb96}
\end{subfigure}
\caption{I/O and CPU breakdown of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}\label{fig:bbd}
\end{figure*}
Unlike the BigBrain incrementation at 25 Processes, the makespan of the BIDS App Example
executing on Optane is longer than local disk (Figure~\ref{fig:bm25}). However, similarly
to the BigBrain Incrementation, tmpfs has a slight lead over local disk and Isilon performs the
worst. The same pattern can be found with 96 process (Figure~\ref{fig:bm96}). When comparing
the executions of 25 and 96 processes, the performance improved by a similar factor on all
storage devices.

@@

As long as all the written data does not exceed the dirty ratio, there is little 
benefit brought by the use of Optane. This is the case with both
our applications writing less than $\SI{1}{\gibi\byte}$ to storage on a system with 
$\SI{768}{\gibi\byte}$ of DRAM and a dirty ratio of 20. However, it is expected 
that as the memory footprints of applications increase to the point where the kernel
would be forced to perform synchronous writes to disk if it were to rely on DRAM alone,
the benefit of using Optane in Memory Mode will become more apparent.

>>>>

\begin{abstract}
    Open-access neuroimaging datasets have reached up to $\SI{}{\peta\byte}$ of data, and
    continue to grow. The ability to leverage the entirety of these datasets are
    limited to a restricted number of labs with both the capacity and infrastructure
    to process the data. Whereas BigData engines have significantly reduced
    application performance penalties with respect to data movement, their applied
    strategies are not necessarily practical within neuroimaging workflows where
    intermediary results may need to be materialized to shared storage for post-processing
    analysis. In this paper we evaluate the performance advantage brought by Intel
    Optane DC persistent memory storage for the processing of large neuroimaging 
    datasets using the two available configurations modes: Memory Mode and App Direct Mode.
    We employ a synthetic algorithm on the $\SI{76}{\gibi\byte}$ \bigbrain
    and a real neuroimaging application on the Consortium for Reliability and Reproducibility (CoRR)
    dataset using 25 and 96 parallel
    processes in both cases. Our results show that performance of applications executing on
    Optane is superior to that of other storage devices, with the exception of DRAM. This is
    the case in both Memory and App Direct Mode and irrespective of the amount of data and 
    parallelism. Furthermore, Optane in App Direct Mode is believed to benefit from the use
    of memory as a cache for writing in cases where output data is significantly smaller than
    available memory. We believe the use of Optane will be beneficial to both neuroimaging applications
    running on HPC or visualization of large, high-resolution images.
\end{abstract}

@@

The server used for the experiments consisted of 12 \SI{64}{\gibi\byte} DRAM devices,
resulting in a total of \SI{768}{\gibi\byte} of DRAM, and 12 \SI{256}{\giga\byte} Intel Optane DC Memory
devices, resulting in total of \SI{3}{\tebi\byte} of Intel Optane DC Memory. The theoretical maximum bandwidth
of DRAM is estimated to be around \SI{275}{\gibi\byte}~/s whereas for Optane, it is
estimated to be around \SI{250}{\gibi\byte}~/s.
Other storage devices included a \SI{240}{\gibi\byte} Micron SATA SSD, of which
only \SI{149}{\gibi\byte} where available for the experiments, and a
\SI{720}{\tebi\byte} shared Dell EMC Isilon cluster. The Isilon cluster is an nsf4 mounted
cluster (10Gb network) consisting of 5 nodes of 36 hard disk drives (HDD). The local disk
was the only storage device set up as a writeback device. Both Isilon and Optane in
App Direct mode were configured as write-through devices, meaning they did not leverage
DRAM as a cache for writes. This was achieved in Optane by configuring the xfs filesystem with
DAX. All storage devices were benchmarked using the following \href{https://github.com/ValHayot/paper-memory-storage/scripts/bench_disks.sh}{script}.
The result of the benchmarks can be seen in Table \ref{table:bandwidths}.

@@

There are some notable differences between the GNU Parallel and Apache Spark implementations.
For instance, in the GNU Parallel implementation, all operations (i.e. reading, incrementing and writing)
are done within the same task. The same task is applied to all the \bigbrain blocks and GNU Parallel is 
provided with all the different command-line calls, processing a subset of them at a time,
depending on the level of parallelism provided. Reading in the GNU Parallel implementation 
is achieved using the popular neuroimaging I/O: Nibabel. When provided with a filename, as is the case
for the GNU Parallel implementation, Nibabel will simply load the header in memory, and memmap the data
using Numpy. This step will be referred to as ``load header''. It is only when the data is actually required
(i.e. during incrementation), that the data will be loaded into memory.

The Apache Spark implementation differs from that of GNU Parallel in that scheduling decisions
were entirely differed to Spark. That is, rather than simply parallelizing the filenames, to be
loaded in a similar fashion to that of GNU Parallel, we chose to load the using Spark's built-in
\texttt{BinaryFiles}, which loads whole files, in binary format, into Spark partitions. As the
data would have been pre-loaded by Spark, ``loading header'' only measures the time to
convert the binary images into Nibabel objects, and ``incrementation'' consists purely
of the time it took to increment the data. Therefore, only the write time would be
comparable between the two implementations, as it is achieved using the same method.
Furthermore, the Spark implementation also differs in that read, increment and write
were separated into three map tasks, which can enable shuffling between the tasks should 
it be determined as necessary by the Spark scheduler.

@@

\subsection{40~$\mu$m \bigbrain incrementation}
\subsubsection{25 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_25cpus.pdf}
    \captionsetup{belowskip=-10pt}
    \caption{Makespan of the \bigbrain Incrementation application using 25 
             processes on the 125 blocks on the 40~$\mu$m image given different
             storage configurations. Anticipated read and write duration was 
             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
             to Eq.~\ref{eq:makespan}. 3 
             repetitions were performed. *Local SSD did not complete the
    writing of the last few blocks due to storage limitations.}\label{fig:makespan-25cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_25cpus.pdf}
    \captionsetup{belowskip=0pt}
    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. 3 repetitions were performed. *Local SSD did not
             complete the writing of the last few blocks due to storage limitations}\label{fig:stacked-25cpus}
\end{figure}
\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572380543-tmpfsAD_1it_25cpus_40bb-1.pdf}
    \caption{DRAM}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572382482-optaneAD_1it_25cpus_40bb-1.pdf}
    \caption{Optane}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572382794-localAD_1it_25cpus_40bb-1.pdf}
    \caption{local SSD*}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572388118-isilonAD_1it_25cpus_40bb-2.pdf}
    \caption{Isilon}\label{fig:gantt25isilon}
\end{subfigure}
    \captionsetup{belowskip=-10pt}
\caption{Gantt charts for each storage device (App Direct Mode) processing 125 blocks of the 40$\mu$m BigBrain using 25 processes. *Some local SSD writes did not complete due to storage limitations.}\label{fig:gantt25}
\end{figure*}

@@

The total task duration breakdowns for each device (Figure~\ref{fig:stacked-25cpus})
shows that, as expected, I/O times vary greatly between devices, with DRAM having the
best total read and write speeds. Optane in Memory Mode is very close to DRAM in speed,
whereas Optane in App Direct mode 7x slower in terms of reading and 23x slower in terms
of writing. Local SSD, on the other hand, exhibits similar write times as DRAM, but is
approximately 96x that DRAM with respect to reads. Header loading for local SSD is also 90x longer
than that of DRAM, however, remains negligible compared to other tasks. Isilon, on the 
other hand, has non negligible header load times, with it being more than 5000 times
slower than that of DRAM. Interestingly enough, the read/increment times on Isilon are
slower than that of the local SSD, although Isilon makes up for its speedier reads with 
writes that are around 184x slower than DRAM.

The App Direct mode Gantt charts (Figure~\ref{fig:gantt25}) also reflect what is observed in Figure~\ref{fig:stacked-25cpus}.
DRAM is significantly faster than all other storage, leading it to barely appear within the Gantt chart.
DRAM was measured to have an average of 17.2 parallel tasks throughout the execution. Optane takes
significantly longer than DRAM and appears to spend a significant amount of time writing. The average
parallelism measured for Optane was 24. Unlike Optane, local SDD spends the vast majority of its time
reading, while spending a negligible amount of time writing. Like Optane, it averages 24 parallel tasks
throughout its execution. Unlike the other three storages, Isilon shows spacing between the loading of the header,
read and increment and write. Isilon was measured to have an average parallelism of 14. Overall, each storage device
execution displayed that some tasks took longer than others.

\subsubsection{96 Processes}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-40bb_96cpus.pdf}
    \captionsetup{belowskip=-10pt}
    \caption{Makespan of the \bigbrain Incrementation application using 96 
             processes on the 125 blocks on the 40~$\mu$m image given different
             storage configurations. Anticipated read and write duration was 
             measured by applying the perceived bandwidths Table~\ref{table:bandwidths}
             to Eq.~\ref{eq:makespan}. 3 
             repetitions were performed. *Local SSD did not complete the
    writing of the last few blocks due to storage limitations.}\label{fig:makespan-96cpus}
\end{figure}
\begin{figure}
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-40bb_96cpus.pdf}
    \captionsetup{belowskip=-10pt}
    \caption{Total read/increment/write breakdown of \bigbrain Incrementation using 25 cpus. 3 repetitions were performed. *Local SSD did not
             complete the writing of the last few blocks due to storage limitations}\label{fig:stacked-96cpus}
\end{figure}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572306855-tmpfs_1it_96cpus_40bb-2.pdf}
    \caption{Memory Mode}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/gantt-1572379747-optaneAD_1it_96cpus_40bb-1.pdf}
    \caption{App Direct Mode}
\end{subfigure}
    \captionsetup{belowskip=-10pt}
\caption{Gantt charts for Optane processing 125 blocks of the 40$\mu$m BigBrain using 96 processes}\label{fig:gantt96}
\end{figure*}

Despite having increased parallelism by a factor of 3.84, we see no visible reduction
in makespan (Figure~\ref{fig:makespan-96cpus}). At 96 cpus, DRAM and Optane in Memory Mode are the only storage
devices that did better than the anticipated I/O, with a makespan duration of
5.7 and 9s, respectively (total read and write estimate for both was measured to be around 39s). While
Optane in App Direct Mode and local SSD both performed worse than the anticipated I/O estimates, Isilon
performed as expected. Variance was found to be high on the local SSD.

The total task duration breakdowns (Figure~\ref{fig:stacked-96cpus}) show that total task durations
have nearly quadrupled. Whereas most devices decreased in performance in both reading and writing,
Isilon only appeared to display a performance decrease with respect to writes. Furthermore, while both
read and writes decreased for Optane in App Direct Mode, read duration was nearly 6x slower with more threads,
while write duration was only about 2x slower. For local SSD and Isilon, there does not appear to 
be any significant difference between App Direct and Memory mode.

When analyzing the average parallelism during the workload, only Optane in App Direct mode
came close to 96 parallel tasks, having an average parallelism of 88 tasks. Optane in Memory mode,
DRAM and local SSD all averaged between 61-66 tasks. Isilon performed the worst with 38 tasks in Memory 
mode and 45 tasks in App Direct mode.

A further look at the Gantt charts between Optane in Memory Mode and App Direct Mode (Figure~\ref{fig:gantt96})
show that both read and write time of tasks are worse in App Direct mode as compare with Memory mode.
While some reads tasks in App Direct mode appear to be of similar duration to the average read task time
of Memory mode, the vast majority of read tasks are significantly slower. Futhermore, not write tasks in App
Direct mode is capable of reaching the performance of equivalent tasks in Memory mode.

@@

The anticipated I/O estimates did not correctly predict the makespan for 25 parallel
processes (Figure~\ref{fig:20mksp25}). Both Optane in Memory Mode and Isilon appear 
to be under the estimates: Optane in Memory Mode is approximately twice as fast as the
estimates whereas Isilon is around 1.1x faster. Optane was 1.8x slower than anticipated.
Once again, Memory and App Direct Mode appear to only affect performance for Optane.

Total task duration breakdowns (Figure~\ref{fig:20total25}) show that for Optane in
App Direct Mode spends significantly more time writing than its Memory Mode counterpart.
In fact, it spends almost 12x more time writing. Isilon also appears to vary slightly
between Memory and App Direct Mode, spending around 1.4x more time reading and incrementing
in App Direct Mode. While the read and write duration appeared to be more-or-less balanced in
Isilon (as was anticipated), Optane in App Direct Mode spent significantly more time writing, whereas the 
inverse was apparent in Memory Mode

Unlike the GNU Parallel executions, the Spark executions appear to be in better accordance with
the I/O estimates (Figure~\ref{fig:20mkspspark25}. Only Optane in App Direct Mode appears
to have significantly exceeded the estimates, by a factor of 2. Total write duration all appear
to be longer in the Spark implementation than with GNU Parallel (Figure~\ref{fig:20totalspark:25}).
Spark's writes are 1.4-1.6x longer with Optane and 1.1x longer with Isilon. Due to differences in
how data is loaded in Spark, it is not possible to comment on how read times are affected. However, it
is possible to note that data conversion (binary string to Numpy array) takes almost twice as 
much time in Optane than it does with Isilon. Furthermore, the act of incrementation can take
up to twice as much time with Isilon.

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_25cpus.pdf}
    \caption{Makespan}\label{fig:20mksp25}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_25cpus.pdf}
    \caption{Total read/increment/write breakdown}\label{fig:20total25}
\end{subfigure}
\captionsetup{belowskip=-10pt}
\caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
25 processes. 3 repetitions were performed.}\label{fig:2025}
\end{figure*}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-spark_20bb_25cpus.pdf}
    \caption{Makespan}\label{fig:20mkspspark25}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-spark-20bb_25cpus.pdf}
    \caption{Total convert/increment/write breadown}\label{fig:20totalspark:25}
\end{subfigure}
\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 25 processes. 3 repetitions wereperformed}\label{fig:20spark25}
\end{figure*}

@@

A very slight performance improvement (1-1.3x faster) can be witnessed in Optane. In 
contrast, Isilon 
experienced a slight performance decrease (1.1x slower). Whereas Isilon makespan appears to match
estimated I/O times, Optane once again has a superior Memory Mode makespan and an inferior App Direct
Mode makespan (Figure~\ref{fig:20mksp96}). As previously observed with 25 processes,
the bulk of the processing time is taken up by writing for Optane in App Direct Mode and 
Isilon in both modes (Figure~\ref{fig:20total96}). Reading takes more time in the case of Optane in Memory mode.

A notable difference in the Spark execution is that Optane in Memory and App Direct Mode
exhibit the same performance (Figure~\ref{20mkspnspark96}). Both of which also performed
significantly worse than the estimated I/O duration. Performance on Isilon, however, did
not really differ. The total task breakdown shows that, as expected, write times are more
significant on Isilon than on Optane and Optane in App Direct Mode spends more time writing
than in Memory Mode (Figure~\ref{20fig:20totalspark96}). When comparing GNU Parallel and 
Spark implementations, Spark spent nearly half as much time writing as GNU Parallel. However,
it was found that Optane spent more time converting the data to Nibabel than GNU Parallel spent
in loading the header.

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-20bb_96cpus.pdf}
    \caption{Makespan}\label{fig:20mksp96}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-20bb_96cpus.pdf}
    \caption{Total read/increment/write breakdown}\label{fig:20total96}
\end{subfigure}
\caption{GNU Parallel incrementation application processing the 20~$\mu$m BigBrain using
96 processes. 3 repetitions were performed.}\label{fig:2096}
\end{figure*}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/makespan-real-spark-20bb_96cpus.pdf}
    \caption{Makespan}\label{fig:20mkspspark96}
\end{subfigure}
\begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bigbrain-incrementation/figures/stacked-real-spark-20bb_96cpus.pdf}
    \caption{Total convert/increment/write breakdown}\label{fig:20totalspark96}
\end{subfigure}
\caption{Spark incrementation application processing the 20~$\mu$m BigBrain using 96 processes. 3 repetitions were performed}\label{fig:20stackedp96}
\end{figure*}
\subsection{BIDS App Example}
\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/25cores.pdf}
    \caption{25 processes}\label{fig:bm25}
\end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/96cores.pdf}
    \caption{96 processes}\label{fig:bm96}
\end{subfigure}
\caption{Makespan of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}
\end{figure*}

\begin{figure*}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/25cores-sum.pdf}
    \caption{25 processes}\label{fig:bb25}
\end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
    \includegraphics[width=\columnwidth]{./experiments/bids-app-example/96cores-sum.pdf}
    \caption{96 processes}\label{fig:bb96}
\end{subfigure}
\caption{I/O and CPU breakdown of BIDS App Example 25 and 96 processes on all storage devices. 3 repetitions were performed}\label{fig:bbd}
\end{figure*}
Unlike the BigBrain incrementation, the makespan of the BIDS App Example
executing on Optane is longer than local disk (Figure~\ref{fig:bm25}). However, similarly
to the BigBrain Incrementation, Optane in Memory Mode performs better than the other storage devices
while Isilon has the longest makespan. The same pattern can be found with 96 process (Figure~\ref{fig:bm96}). When comparing
the executions of 25 and 96 processes, the performance improved by a similar factor on all
storage devices.

@@

The use of Optane in Memory Mode would enable more dirty data to be written to available memory
rather than directly to slower storage devices. In our 40~$\mu$ \bigbrain experiments, it was found
that read times were less important in Isilon than the local SSD (Figures~\ref{fig:stacked-25cpus}
and~\ref{fig:stacked-96cpus}). Should Isilon have been given the opportunity to benefit from a 
writeback cache like the SSD, it is believed that performance on Isilon could have been superior 
to that of the SSD. Having Optane as a node-local burst buffer to shared network storage 
may prove to be very beneficial in the case of Big Data neuroimaging pipelines running on
High Performance Computing clusters.
